AttCRISPR : an spacetime interpretable model for sgRNA efficiency prediction

### 2020-10-03 Paper/main.tex
10-3 paper

\begin{figure}[!tpb]%figure4
    \centerline{\includegraphics[scale=0.15]{CNNv2.png}}
    \caption{The architecture of temporal domain method in AttCRISPR. 
    The input of the method is embedding sgRNA sequence, a $21\times e$ embedding matrix, where $e$ is the dimension of a nucleotide embedding vector. 
    }\label{fig:04}
\end{figure}
As Figure~4\vphantom{\ref{fig:04}} show, temporal domain part of AttCRISPR relies on the recurrent neural network (or its variant). 
As previously mentioned, we mapping each one-hot vector into a dense real-valued high-dimensional space follow the Equation~(\ref{eq:08}), which generates the embedded matrix $X_e$.
And we regard $X_{e}$ as a sequential data, or temporal data.
Recurrent neural network (or its variant) has showed outstanding performance in the tasks with temporal data (for instance, natural language processing, sequential recommendation). 
That's why we prefer to use it to extract potential temporal features. 
To be precise, we prefer the architecture of encoder-decoder which has been proven to be effective in Seq2Seq task. 
Two main differences we have to face are that sgRNA is not a natural language in the traditional sense, and we don't have to translate it to other sequence. 
To accommodate them, the embedded matrix $X_e$ is used as input of both the encoder and decoder, and the sequence of decoder is to build the first-order preference of sgRNA sequence $\tilde{A}$. 
As mentioned above, the predicted value $y$ should satisfy the following equation
\begin{equation}
y=\sum_i^lW_i\tilde{A}_i\cdot {X_e}_i\label{eq:12}\vspace*{-10pt}
\end{equation}\\
where $W$ denotes a non-negative weight vector acquired through a full connection layer, $W_i$ denotes the $i$-th dimension of the vector $W$. 
Equation~(\ref{eq:12}) is the same formally as Equation~(\ref{eq:04}). 
On this basis, we apply the idea of attention mechanism which has been widely used in natural language processing tasks to AttCRISPR in the method of temporal domain, 
and name it Temporal Attention Module \citep{vaswani2017attention,luong2015effective,bahdanau2014neural}.
As the idea of \citeauthor{vaswani2017attention}, Temporal Attention Module satisfies the following equation
\begin{equation}
Attention(Q,K,V)=align(Q,K)V\label{eq:12}\vspace*{-10pt}
\end{equation}\\
where the $align(\cdot)$ is put forward by \citeauthor{luong2015effective}. 
$Q$, $K$, $V$ are queries, keys and values matrix accordingly in the paper of \citeauthor{vaswani2017attention}. 
Specifically, in our attention module they are calculated by the following equation
\begin{equation}
K_i=Encoder({X_e}_i,\theta_E,K_{i-1})\label{eq:13}
\end{equation}
\begin{equation}
Q_i=Decoder({X_e}_i,\theta_D,Q_{i-1})\label{eq:14}
\end{equation}
\begin{equation}
V=K\label{eq:15}\vspace*{-10pt}
\end{equation}\\
where vector $K_i$, $Q_i$ denotes the $i$-th row of the matrix $K$ and $Q$ accordingly, $Encoder(\cdot)$ and $Decoder(\cdot)$ are independent GRU units, $\theta_E$ and $\theta_D$ denote all the related parameters of GRU networks accordingly. 
In the actual implementation, we apply the bidirectional GRU networks for better performance, and for the sake of conciseness, we show a conventional GRU network here. 
The function $align()$ is as follows
\begin{equation}
B=align(Q,K)\label{eq:16}
\end{equation}
\begin{equation}
B_i=softmax(Q_iK^T)\otimes G_i\label{eq:17}
\end{equation}
\begin{equation}
G_{ij}=\left\{\begin{matrix}
exp(\frac{(i-j)^2}{-2\sigma})&,\left | i-j \right |\leqslant \sigma
\\ 0&,\left | i-j \right |>  \sigma
\end{matrix}\right.\label{eq:18}\vspace*{-10pt}
\end{equation}\\
where the matrix $B\in\mathbb{R}^{l\times l}$ is the second-order preference we need, and vector $B_i$ denotes the $i$-th row of the matrix $B$. 
In our belief, the $j$-th dimension of $B_i$, denoted as $B_{ij}$, can reveal the effect of the base at position $j$ on position $i$ in a biological sense.

>>>>

\begin{figure}[!tpb]%figure4
    \centerline{\includegraphics[scale=0.15]{CNNv2.png}}
    \caption{The architecture of temporal domain method in AttCRISPR. 
    The input of the method is embedding sgRNA sequence, a $21\times e$ embedding matrix, where $e$ is the dimension of a nucleotide embedding vector. 
    }\label{fig:04}
\end{figure}
As Figure~4\vphantom{\ref{fig:04}} show, temporal domain part of AttCRISPR relies on the recurrent neural network (or its variant). 
As previously mentioned, we mapping each one-hot vector into a dense real-valued high-dimensional space follow the Equation~(\ref{eq:08}), which generates the embedded matrix $X_e$.
And we regard $X_{e}$ as a sequential data, or temporal data.
Recurrent neural network (or its variant) has showed outstanding performance in the tasks with temporal data (for instance, natural language processing, sequential recommendation). 
That's why we prefer to use it to extract potential temporal features. 
To be precise, we prefer the architecture of encoder-decoder which has been proven to be effective in Seq2Seq task. 
Two main differences we have to face are that sgRNA is not a natural language in the traditional sense, and we don't have to translate it to other sequence. 
To accommodate them, the embedded matrix $X_e$ is used as input of both the encoder and decoder, and the sequence of decoder is to build the first-order preference of sgRNA sequence $\tilde{A}$. 
As mentioned above, the predicted value $y$ should satisfy the following equation
y=\sum_i^lW_i\tilde{A}_i\cdot {X_e}_i\label{eq:12}\vspace*{-10pt}
where $W$ denotes a non-negative weight vector acquired through a full connection layer, $W_i$ denotes the $i$-th dimension of the vector $W$. 
Equation~(\ref{eq:12}) is the same formally as Equation~(\ref{eq:04}). 
On this basis, we apply the idea of attention mechanism which has been widely used in natural language processing tasks to AttCRISPR in the method of temporal domain, 
and name it Temporal Attention Module \citep{vaswani2017attention,luong2015effective,bahdanau2014neural}.
As the idea of \citeauthor{vaswani2017attention}, Temporal Attention Module satisfies the following equation
Attention(Q,K,V)=align(Q,K)V\label{eq:12}\vspace*{-10pt}
where the $align(\cdot)$ is put forward by \citeauthor{luong2015effective}. 
$Q$, $K$, $V$ are queries, keys and values matrix accordingly in the paper of \citeauthor{vaswani2017attention}. 
Specifically, in our attention module they are calculated by the following equation
K_i=Encoder({X_e}_i,\theta_E,K_{i-1})\label{eq:13}
Q_i=Decoder({X_e}_i,\theta_D,Q_{i-1})\label{eq:14}
V=K\label{eq:15}\vspace*{-10pt}
where vector $K_i$, $Q_i$ denotes the $i$-th row of the matrix $K$ and $Q$ accordingly, $Encoder(\cdot)$ and $Decoder(\cdot)$ are independent GRU units, $\theta_E$ and $\theta_D$ denote all the related parameters of GRU networks accordingly. 
In the actual implementation, we apply the bidirectional GRU networks for better performance, and for the sake of conciseness, we show a conventional GRU network here. 
The function $align()$ is as follows
B=align(Q,K)\label{eq:16}
B_i=softmax(Q_iK^T)\otimes G_i\label{eq:17}
G_{ij}=\left\{\begin{matrix}
exp(\frac{(i-j)^2}{-2\sigma})&,\left | i-j \right |\leqslant \sigma
\\ 0&,\left | i-j \right |>  \sigma
\end{matrix}\right.\label{eq:18}\vspace*{-10pt}
where the matrix $B\in\mathbb{R}^{l\times l}$ is the second-order preference we need, and vector $B_i$ denotes the $i$-th row of the matrix $B$. 
In our belief, the $j$-th dimension of $B_i$, denoted as $B_{ij}$, can reveal the effect of the base at position $j$ on position $i$ in a biological sense. 
G is the damping matrix base on Gaussian function. 
Since a simple belief that the closer the base is to the $i$-th position, the more it affects the $i$-th position normally, we use the damping matrix G to constrain the network learning. 
$\sigma$ represents a threshold of length, any base over this length from the position $i$ is not considered to be affected.
